\chapter{Monsters}

\begin{multicols}{2}

This chapter lists monsters and other creatures that the
players might encounter.

The monster entries list out the monster's statistics and
describe the monster.

Armor Class is the monster's armor class.

Hit Dice is the number of dice the Game Master roles to
determine the monster's Hit Points when it appears.

Move is the movement the monster is capable of in a single
round.

Attacks is a list of the different attacks a monster can
make.

No. Appearing is the number of monsters of that type that
the players run across in a single encounter.

Treasure Type is the treasure that the monster has on it or
in a nearby location.

Intelligence is the general intelligence of the monster. This
follows the normal 3-18 range.

Alignment is the alignment that the monster is attuned to.

\section{Ganter}

\textbf{Armor Class:} 13

\textbf{Hit Dice:} 1

\textbf{Move:} 30'

\textbf{Attacks:} Claws

\textbf{No. Appearing:} 2-8 (2d4)

\textbf{Treasure Type:} Meager

\textbf{Intelligence:} 7

\textbf{Alignment:} Void

Ganters are vicious predators. They are humanoid, but with large
elongated heads and reptilian skin. They have three eyes, and on
each hand, three clawed fingers. Though sentient, their primary
goal is to devour all sentient living things. Ganter society is
organized on the principle that death is the natural state of
all things, and they happily practice their beliefs by killing
and eating everyone they come across. They are a joyful race and
honestly believe they are doing everyone a favor.

\section{Goblin}

\textbf{Armor Class:} 11

\textbf{Hit Dice:} 1

\textbf{Move:} 30'

\textbf{Attacks:} Short sword, spear, or sling

\textbf{No. Appearing:} 1-4 (1d4)

\textbf{Treasure Type:} Meager

\textbf{Intelligence:} 7

\textbf{Alignment:} Chaos

Goblins are small, foul-smelling creatures with reedy limbs,
pointed ears, and crooked noses. Their yellow eyes look greedily
upon everything they see. The driving force behind a goblin's
decisions is hunger first and greed second. They are cowardly
creatures.

\section{Hobgoblin}

\textbf{Armor Class:} 15

\textbf{Hit Dice:} 1

\textbf{Move:} 30'

\textbf{Attacks:} Spear, trident, or throwing axe

\textbf{No. Appearing:} 1-4 (1d4)

\textbf{Treasure Type:} Raider

\textbf{Intelligence:} 10

\textbf{Alignment:} Chaos

Hobgoblins are larger, more intelligent, and more vicious
cousins to goblins. They have yellow or sand-colored skin,
bright red eyes, and stand around six feet tall. Where
goblins are cowardly, hobgoblins are devious. While they
won't shy from a fight, they prefer to lay traps and set
ambushes.

\section{Orc}

\textbf{Armor Class:} 14

\textbf{Hit Dice:} 1

\textbf{Move:} 30'

\textbf{Attacks:} Short sword, hand axe, or short bow

\textbf{No. Appearing:} 1-4 (1d4)

\textbf{Treasure Type:} Meager

\textbf{Intelligence:} 8

\textbf{Alignment:} Chaos

Orcs are brutish creatures with pointy ears and green skin. They
have short tusks protruding from their lower jaw and a slanted
brow. Large and violent, orcs are prone to raiding and pillaging.
They have thrived in the chaos that followed the Breaking. Some
say there are orcish cities in the wilderness where bloodshed is
the only law.

\section{Orthrus}

\textbf{Armor Class:} 14

\textbf{Hit Dice:} 3

\textbf{Move:} 60'

\textbf{Attacks:} Two bites

\textbf{No. Appearing:} 1

\textbf{Treasure Type:} None

\textbf{Intelligence:} 5

\textbf{Alignment:} Destiny

An orthrus is a two-headed hound commonly found living with
giants. It stands six feet at the shoulder and comes in a
variety of mud colors. Orthruses are loyal companions and
more intelligent than most beasts, though their two heads
sometimes compete with each other over affections.

\end{multicols}
